---
title: "Visualization of Controls over DIV"
author: "Amy Carpenter"
date: "5/19/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(stringi)
```

# Purpose
Create a visualization of the parameter values over DIV time

**Ugh, this is not doing what I want it to do with looping over the plots.

I'll just use an R script to make some pdf's instead


# Load data 

```{r}
load('lvl0_snapshots/mea_nfa_lvl0_extra_cols_2021-05-05.RData')

# Make the artificial point at DIV = 2 clear
add.div2 <- mea_nfa_lvl0[grepl('DIV12',acnm)]
add.div2[, acnm := sub('DIV12','DIV2',acnm)]
add.div2[, rval := NA_real_]
mea_nfa_lvl0 <- rbind(mea_nfa_lvl0, add.div2)
rm(add.div2)

# Add DIV
mea_nfa_lvl0[, DIV_char := stri_extract(acnm, regex = 'DIV[0-9]*')]
mea_nfa_lvl0[, DIV := as.numeric(stri_extract(DIV_char, regex = '[0-9]{1,2}'))]

# Note where NA rvals set to 0 for AUC calculation
mea_nfa_lvl0[, plot_rval := ifelse(!is.na(rval), rval, 0)]
mea_nfa_lvl0[, added_0 := as.numeric(is.na(rval))]
mea_nfa_lvl0$added_0 <- factor(mea_nfa_lvl0$added_0, levels = c('0','1'), ordered = T)

```


# Make plots

```{r}
controldat <- mea_nfa_lvl0[wllt == 'n' & wllq == 1]
rm(mea_nfa_lvl0)
controldat[, culture := sub('_.*$','',apid)]
controldat[, well_id := paste0(apid, '_', rowi,'_',coli)]
controldat[, plate_id := frank(apid, ties.method='dense'), by = .(culture)]
controldat[, plate_char := paste0('Plate',plate_id)]
use.cultures <- sample(unique(controldat[plate_id >= 6, culture]), size = 5)

plotdat <- controldat[culture %in% use.cultures & grepl('firing_rate_mean_DIV',acnm)]

ggplot(plotdat[culture %in% use.cultures], aes(x = DIV, y = plot_rval)) +
  geom_point(aes(color = added_0))+
  geom_line(aes(group = well_id))+
  facet_grid(rows = vars(culture), 
             cols = vars(plate_char))+
  scale_color_manual(values = c('0' = 'black',
                                '1' = 'red'))+
  scale_x_continuous(breaks = c(2,5,7,9,12))+
  # scale_y_continuous(limits = quantile(plotdat$plot_rval, c(0, 0.99))) + # don't let outliers throw off the scale
  ggtitle('Mean Firing Rate Over DIV')

```



# Mean Burst Duration
```{r eval = FALSE}
acnm_sub <- 'CCTE_Shafer_MEA_dev_burst_duration_mean_DIV'
acnm_sub <- 'CCTE_Shafer_MEA_dev_per_burst_interspike_interval_DIV'
plotdat <- controldat[culture %in% use.cultures & grepl(acnm_sub, acnm)]

ggplot(plotdat[culture %in% use.cultures], aes(x = DIV, y = plot_rval)) +
  geom_point(aes(color = added_0))+
  geom_line(aes(group = well_id))+
  facet_grid(rows = vars(culture), 
             cols = vars(plate_char),
             scales = 'free_y',
             switch = 'y')+
  scale_color_manual(values = c('0' = 'black',
                                '1' = 'red'))+
  scale_x_continuous(breaks = c(2,5,7,9,12))+
  # scale_y_continuous(limits = quantile(plotdat$plot_rval, c(0, 0.99))) + # don't let outliers throw off the scale
  ggtitle(paste0(sub('_DIV','',acnm_sub),' over DIV'))
```


# Use loop function

## Create plots
```{r}
acnms <- unique(controldat[is.na(DIV) & !grepl('(LDH)|(AB)',acnm), acnm])[1:3]
plotdat <- controldat[culture %in% use.cultures]

plot.object <- list()

for (acnm_sub in paste0(acnms,'_DIV')) {
  
  p <- ggplot(plotdat[grepl(acnm_sub, acnm)], aes(x = DIV, y = plot_rval)) +
    geom_point(aes(color = added_0))+
    geom_line(aes(group = well_id))+
    facet_grid(rows = vars(culture), 
               cols = vars(plate_char),
               scales = 'free_y',
               switch = 'y')+
    scale_color_manual(values = c('0' = 'black',
                                  '1' = 'red'))+
    scale_x_continuous(breaks = c(2,5,7,9,12))+
    # scale_y_continuous(limits = quantile(plotdat$plot_rval, c(0, 0.99))) + # don't let outliers throw off the scale
    ggtitle(paste0(sub('_DIV','',acnm_sub),' over DIV'))+
    theme(legend.position = 'top')+
    ylab('rval')
  plot.object[[acnm_sub]] <- p
}

```


## Testing

```{r echo = FALSE, eval = TRUE}
out <- lapply(seq_along(plot.object), function(i) {
  a1 <- knitr::knit_expand(text = sprintf("### %s\n", names(plot.object)[i])) # tab header, auto extracts names of `hcs`
  a2 <- knitr::knit_expand(text = "\n```{r}") # start r chunk
  a3 <- knitr::knit_expand(text = sprintf("\nplot.object[[%s]]",i)) # extract graphs by "writing" out `hcs[[1]]`, `hcs[[2]]` etc. to be rendered later
  a4 <- knitr::knit_expand(text = "\n```\n") # end r chunk
  paste(a1, a2, a3, a4, collapse = '\n') # collapse together all lines with newline separator
})
cat(unlist(out), sep = '')
```

Want to use this, but it is not working when out includes any type of plot
`r knitr::knit(text = paste(out, collapse = '\n'))`


### CCTE_Shafer_MEA_dev_burst_rate_DIV
 
```{r} 
print(plot.object[[1]])
```
### CCTE_Shafer_MEA_dev_interburst_interval_mean_DIV
 
```{r} 
print(plot.object[[2]] )
```
### CCTE_Shafer_MEA_dev_burst_duration_mean_DIV
 
```{r} 
print(plot.object[[3]] )